---
title: "synthetic_data"
author: "DM 21"
date: "2024-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(readr)
library(RSQLite)
library(dplyr)
library(generator)
library(charlatan)
library(digest)
library(stringi)
library(chron)
library(DBI)
library(lubridate)
library(scales)
library(ggplot2)
```

# Generate Synthetic data
In this session, several R packages and address datasets have been utilised to generate synthetic data that fits the database schema we created earlier.

## Personal data generation
To generate personal data, including addresses and phone numbers, real UK postcode data has been used, which contains information about cities and counties. Additionally, random street names were extracted from the 'street_name.csv' dataset to ensure synthetic address data. For phone numbers, British random phone numbers starting with '07' and nine digits in length were generated.
```{r}
address_db <- read.csv("postcodes.csv", na.strings = c("", "NA"), fileEncoding = "UTF-8")
address_db <- address_db[, c(1,6,9)]
address_db <- na.omit(address_db)
street_db <- read.csv("street_name.csv")
postcode <- address_db$postcode
st_name <- street_db$Street.Name

# Function to generate phone numbers starting with 07
phone_number_custom <- function(n) {
  phone_numbers <- stri_rand_strings(n, pattern = "[0-9]", length = 9)
  phone_numbers <- paste0("07", phone_numbers)
  return(phone_numbers)
}
```

## Generating datasets for tables

### Customer data
We set a specific date range for cust_reg_date, ensuring that this range falls later than the customer's birthday (cust_birth_date). Customer names and passwords were generated using the 'generator' package, with the names split into first names and last names. Additionally, we ensured that cust_password was encrypted. Zip codes and addresses were generated using the dataset mentioned above. 
```{r}
n <- 200
set.seed(1)
membership_type <- c("SILVER", "BRONZE", "GOLD", "PLATINUM")

customer <- data.frame(
  cust_id = seq(n),
  cust_reg_date = r_date_of_births(n, start = as.Date("2018-01-01"), end = Sys.Date()),
  cust_full_name = ch_name(n, locale = "en_GB", messy = FALSE),
  cust_email = r_email_addresses(n),
  cust_password = r_national_identification_numbers(n),
  cust_phone = phone_number_custom(n),
  postcode = sample(x = postcode, size = n, replace = TRUE),
  cust_street_address = paste0(sample(x = st_name, size = n, replace = TRUE), " ", sample(1:99, n, replace = TRUE)),
  membership = sample(x = membership_type, size = n, replace = TRUE),
  cust_birth_date = r_date_of_births(n, start = as.Date("1970-01-01"), end = as.Date("2010-01-01")),
  cust_referral = sample(0:n, size = n, replace = TRUE),
  stringsAsFactors = FALSE
)

# Split the first name and last name
customer$cust_first_name <- sapply(strsplit(as.character(customer$cust_full_name), " "), function(x) paste(x[-length(x)], collapse = " "))
customer$cust_last_name <- sapply(strsplit(as.character(customer$cust_full_name), " "), function(x) tail(x, 1))

# Customer password
customer$cust_password <- digest(customer$cust_password, algo = "md5", serialize = FALSE)

# Customer address
customer <- merge(customer, address_db, by = "postcode", all.x = TRUE)
customer$cust_zipcode <- customer$postcode
customer$cust_city <- customer$town
customer$cust_county <- customer$country_string

# Select customer data
customer <- select(customer, cust_id, cust_reg_date, cust_last_name, cust_first_name, cust_email, cust_password, cust_phone, cust_zipcode, cust_street_address, cust_city, cust_county, membership, cust_birth_date, cust_referral)

# Generate CSV data
write.csv(customer, file = "customer.csv", row.names = FALSE)

# Print
print(customer)
```

### Product data
In the product data, 30 entries are distributed among 10 distinct categories. To ensure consistency, the structure of the 'model_name' attribute across all products has been standardised. The 'cost_price' of each product is randomly assigned within a range of 85% to 90% of the 'selling_price', maintaining variability while adhering to predetermined criteria.
```{r}
n <- 30
set.seed(1)

# Generate product data frame
product <- data.frame(
  product_id = 1:n,
  category_name = sample(c('guitar', 'drum', 'base', 'keyboards', 'pianos', 'saxophones', 'flute', 'violin', 'cello', 'harp', 'chimes', 'french horn', 'trumpet', 'bagpipe', 'microphone', 'cable', 'connector', 'lightning', 'headphone'), n, replace = TRUE),
  model_name = sapply(1:30, function(x) paste0(sample(letters, 2, replace = TRUE), sample(0:9, 2, replace = TRUE), collapse = '')),
  selling_price = NA,  # Initialize selling_price column
  stock_qty = sample(30:300, n, replace = TRUE)
)

# Set selling price based on category_name
product$selling_price <- ifelse(product$category_name == "harp", 
                                sample(1100:3000, n, replace = TRUE), 
                                ifelse(product$category_name %in% c("guitar", "base", "flute", "violin", "cello", "chimes", "french horn", "trumpet", "bagpipe"), 
                                       sample(1000:2599, n, replace = TRUE), 
                                       sample(50:999, n, replace = TRUE)))

# Add cost_price column
product$cost_price <- ifelse(product$category_name == 'pianos', product$selling_price * 0.90,
                             ifelse(product$category_name %in% c('guitar', 'drum', 'base', 'keyboards', 'saxophones', 'flute', 'violin', 'cello', 'harp', 'chimes', 'french horn', 'trumpet', 'bagpipe'), product$selling_price * 0.85,
                                    product$selling_price * 0.90))

# Save product.csv
write.csv(product, file = "product.csv", row.names = FALSE)

# Print
print(product)
```

### Promotion data
We set expiration dates in 2024 for all promotion.
```{r}
n <- 30

# Set seed
set.seed(2)

# Generate promotion IDs
promotion_id <- seq(n)

# Generate promotion rate convert to percentage
promotion_name <- sample(c('BirthdayBonus', 'VIP', 'Referral', 'SpringSale', 'BackTOSchool', 'MerryChristmas', 'WelcomeNewYear', 'ValentineValue', 'NewcommerWelcome', 'EasterSurprise'), 30, replace = TRUE)

# Combine data
promotion <- data.frame(promotion_id, promotion_name = promotion_name)

# Save the data frame to a CSV file
write.csv(promotion, "promotion.csv", row.names = FALSE)

# Print
print(promotion)
```


### Voucher Data

The generated voucher rates are in decimal.
```{r}
n <- 30

# Set seed
set.seed(2)

# Generate promotion IDs
voucher_code <- paste0("UK", sprintf("%02d", 1:30))

# Generate promotion rate convert to percentage
voucher_rate <- sample(c('0.05', '0.10', '0.15', '0.20', '0.25', '0.30'), 30, replace = TRUE)

# Combine data
voucher <- data.frame(
  voucher_code = voucher_code,
  voucher_rate = voucher_rate,
  promotion_id = sample(promotion$promotion_id, n, replace = TRUE),
  stringsAsFactors = FALSE
)

# Save the data frame to a CSV file
write.csv(voucher, "voucher.csv", row.names = FALSE)

# Print
print(voucher)
```

### Ad data
The company utilises a total of 6 types of media for advertisement, with each ad_name created based on the ad_type. We ensured that the end_date of each ad is later than its start_date. Additionally, to reflect industry norms, we enforced constraints such as ensuring that impression > click > action, with specific conditions such as click < impression * 0.1. 
```{r}
n <- 200

# Define the function to generate synthetic data with specified conditions
generate_synthetic_data <- function(n) {
  media_type <- c("Website", "Facebook", "Instagram", "Tiktok", "Google", "Flyer")
  ad_type <- c("Discount", "Cut Off", "Promotion", "Bonus", "Freebies")
  
  ad_data <- data.frame(
    ad_id = seq(n),
    media_name = sample(x = media_type, size = n, replace = TRUE),
    ad_type = paste0(sample(x = ad_type, size = n, replace = TRUE)),
    start_date = as.Date(sample(seq(as.Date("2018-01-01"), Sys.Date(), by = "day"), n, replace = TRUE)),
    impression = sample(100:1000, n, replace = TRUE),
    stringsAsFactors = FALSE
  )
  
  ad_data$end_date <- ad_data$start_date + sample(1:90, n, replace = TRUE)
  ad_data$impression <- sample(1000:50000, n, replace = TRUE)
  ad_data$cost <- sample(10:1500, n, replace = TRUE)
  ad_data$revenue <- ad_data$cost * sample(seq(0.1, 5, by = 0.01), n, replace = TRUE)
  ad_data$click <- round(ad_data$impression * runif(n, 0.05, 0.1))
  ad_data$action <- round(ad_data$click * runif(n, 0.05, 0.2))
  
  # Filter out rows based on conditions
  ad <- ad_data %>%
    filter(impression >= 1000,
           click < impression * 0.1,
           action < click * 0.2,
           cost >= click * 0.3,
           cost <= click * 2,
           revenue >= cost * 0.1,
           revenue <= cost * 5)
  
  return(ad_data)
}


# Generate synthetic data
n <- 200
ad <- generate_synthetic_data(n)

# Write data to CSV
write.csv(ad, file = "ad.csv", row.names = FALSE)

# Print
print(ad)
```


### Supplier data
The company purchases the products from 10 contracted suppliers, and suppliers' information including phone number and address were randomly generated in British style.
```{r}
n <- 10
s_phone <- phone_number_custom(n)
s_email <- r_email_addresses(n)

supplier <- data.frame(
  s_id = seq(n),
  s_name = c('Harmony Haven Instruments', 'Melody Masters Music Emporium', 'Rhythmic Resonance Supplies', 'Crescendo Corner', 'Sonic Spectrum Distributors', 'Tempo Treasures Inc.', 'Aria Attic Instruments', 'Virtuoso Ventures', 'Serenade Suppliers', 'Allegro Accessories Outlet'),
  postcode = sample(x = postcode, size = n, replace = TRUE),
  s_street_address = paste0(sample(x = st_name, size = n, replace = TRUE), " ", sample(1:10, n, replace = TRUE)),
  s_phone = s_phone,
  s_email = s_email
)

supplier <- merge(supplier, address_db, by = "postcode", all.x = TRUE)
supplier$s_zipcode <- supplier$postcode
supplier$s_city <- supplier$town
supplier$s_county <- supplier$country_string

supplier <- select(supplier, s_id, s_name, s_zipcode, s_street_address, s_city, s_county, s_phone, s_email)

# Display the generated data
print(supplier)

# Save supplier.csv
write.csv(supplier, file = "supplier.csv", row.names = FALSE)
```

### Warehouse data
The company operates one warehouse for each region of England, Wales, and Scotland. The warehouse names (w_name) are derived from the names of the respective regions (counties). For example, warehouse 'Eng01' corresponds to the region of England and has address information based on England. 
```{r}
n <- 3

# Create warehouse data frame with three rows
warehouse <- data.frame(
  w_id = seq(n),
  w_name = c('Eng01', 'Wales01', 'Scot01'),
  w_street_address = paste0(sample(x = st_name, size = n, replace = TRUE), " ", sample(1:10, n, replace = TRUE)),
  w_county = c('England', 'Wales', 'Scotland'),
  stringsAsFactors = FALSE  # Avoids creating factors for character variables
)

# Sample postcodes for each region
sampled_postcodes <- sapply(warehouse$w_county, function(county) {
  subset_postcodes <- address_db$postcode[address_db$country_string == county]
  if (length(subset_postcodes) > 0) {
    return(sample(subset_postcodes, 1))
  } else {
    return(NA)  # Handle the case where there are no postcodes for the specified county
  }
})

# Add sampled postcodes to the warehouse data
warehouse$postcode <- sampled_postcodes

# Assuming address_db is a data frame or tibble with columns w_zipcode, w_city, etc.
# Merge with address_db
warehouse <- merge(warehouse, address_db, by = "postcode", all.x = TRUE)
warehouse$w_zipcode <- warehouse$postcode
warehouse$w_city <- warehouse$town

# Select relevant columns
warehouse <- warehouse %>%
  select(w_id, w_name, w_zipcode, w_street_address, w_city, w_county)

# Save warehouse.csv
write.csv(warehouse, file = "warehouse.csv", row.names = FALSE)

# Print the warehouse data
print(warehouse)
```

### Promote data
The 'promote' table establishes a many-to-many relationship between advertisements and products. Each row from the previously generated product and ad data was brought into the 'promote' table to reflect this relationship.
```{r}
n <- 200

# Create promote data frame
promote <- data.frame(ad_id = sample(ad$ad_id, n, replace = TRUE),
                      product_id = sample(product$product_id, n, replace = TRUE),
                      stringsAsFactors = FALSE)

# Save promote.csv
write.csv(promote, file = "promote.csv", row.names = FALSE)

# Print the promote data
print(promote)

```


### Stock table
The company assigns a unique sku to each product_id, ensuring that products with the same product ID are stored in the same location within each warehouse.
```{r}
# Set seed for reproducibility
set.seed(1)

# Create stock data frame
stock <- data.frame(
  sku = 1:30,  # Assuming 30 rows as there are 30 products
  product_id = sample(product$product_id, 30, replace = TRUE),
  w_id = sample(warehouse$w_id, 30, replace = TRUE)
)

# Display the generated stock data
print(stock)

# Save stock.csv
write.csv(stock, file = "stock.csv", row.names = FALSE)
```

### Order data
The order data was generated within a certain range of dates, with additional data concentrated in specific time periods to capture monthly trends. We ensured that cust_id, product_id, and voucher_code were referenced from existing data generated earlier. Additionally, for orders with the same order_id, the payment_method was standardized as only one payment method can be used for each order.
```{r}
library(dplyr)

n <- 500

# Generate order IDs
order_id <- seq(n)

# Generate customer order dates
start_date <- as.Date("2022-01-01")
end_date <- as.Date("2024-12-31")

# Generate more orders in March, September, and December
extra_dates <- c(
  sample(seq(as.Date("2024-03-01"), as.Date("2024-03-31"), by = "day"), size = 30, replace = TRUE),
  sample(seq(as.Date("2024-09-01"), as.Date("2024-09-30"), by = "day"), size = 30, replace = TRUE),
  sample(seq(as.Date("2024-12-01"), as.Date("2024-12-31"), by = "day"), size = 50, replace = TRUE),
  sample(seq(as.Date("2023-03-01"), as.Date("2023-03-31"), by = "day"), size = 25, replace = TRUE),
  sample(seq(as.Date("2023-09-01"), as.Date("2023-09-30"), by = "day"), size = 20, replace = TRUE),
  sample(seq(as.Date("2023-12-01"), as.Date("2023-12-31"), by = "day"), size = 40, replace = TRUE),
  sample(seq(as.Date("2022-03-01"), as.Date("2022-03-31"), by = "day"), size = 20, replace = TRUE),
  sample(seq(as.Date("2022-09-01"), as.Date("2022-09-30"), by = "day"), size = 15, replace = TRUE),
  sample(seq(as.Date("2022-12-01"), as.Date("2022-12-31"), by = "day"), size = 30, replace = TRUE)
)

order_date <- c(sample(seq(start_date, end_date, by = "day"), size = 240), extra_dates)

# Generate quantities as three-digit numbers that are multiples of 5
quantity <- sample(1:5, size = n, replace = TRUE)

# Generate customer IDs
cust_id <- sample(customer$cust_id, n, replace = TRUE)

# Generate product IDs
product_id <- sample(product$product_id, n, replace = TRUE)

# Generate payment methods (assuming 1 = Credit Card, 2 = Debit Card, 3 = Cash)
# Generate a consistent payment method for each unique order_id, cust_id, and product_id
payment_methods <- c("Credit Card", "Debit Card", "Paypal")
payment_method <- rep(payment_methods, length.out = n)

# Generate promotion IDs
voucher_code <- sample(voucher$voucher_code, n, replace = TRUE)

# Combine data
order <- data.frame(
  order_id = order_id,
  order_date = order_date,
  quantity = quantity,
  cust_id = cust_id,
  product_id = product_id,
  payment_method = payment_method,
  voucher_code = voucher_code
)

# Save the data frame to a CSV file
write.csv(order, "order.csv", row.names = FALSE)

# View first few rows
head(order)

```

### Sell data
To establish the relationship between suppliers and products, we ensured that the supplier IDs (s_id) and product IDs (product_id) referenced existing data generated earlier.
```{r}
n<-50

# Set seed (optional)
set.seed(123)

# Combine data
sell <- data.frame(
  s_id = sample(supplier$s_id, n, replace = TRUE),
  product_id = sample(product$product_id, n, replace = TRUE)
)

# Save the data frame to a CSV file
write.csv(sell, "sell.csv", row.names = FALSE)

# View first few rows
head(sell)

```
## Read the csv file
```{r}
customer <- readr::read_csv("customer.csv", col_types=cols(
  cust_reg_date = col_character(),
  cust_birth_date = col_character()
)) 
ad <- readr::read_csv("ad.csv", col_types=cols(
  start_date = col_character(),
  end_date = col_character()
))
warehouse <- readr::read_csv("warehouse.csv", col_types=cols())
voucher <- readr::read_csv("voucher.csv", col_types=cols())
promotion <- readr::read_csv("promotion.csv", col_types=cols())
product <- readr::read_csv("product.csv", col_types=cols()) 
order <- readr::read_csv("order.csv", col_types=cols(order_date = col_character()))
sell <- readr::read_csv("sell.csv", col_types=cols())
supplier <- readr::read_csv("supplier.csv", col_types=cols()) 
promote <- readr::read_csv("promote.csv", col_types=cols()) 
stock <- readr::read_csv("stock.csv", col_types=cols())
```

## Connect database with data
Finally, we tested if all data fit into database before automate workflow.
```{r}
# Establish a connection to your SQLite database
con <- dbConnect(RSQLite::SQLite(), dbname = "e_commerce.db")

# Write the data frame to the database

dbWriteTable(con, "customer", customer, append = TRUE, row.names = FALSE)
dbWriteTable(con, "ad", ad, append = TRUE, row.names = FALSE)
dbWriteTable(con, "warehouse", warehouse, append = TRUE, row.names = FALSE)
dbWriteTable(con, "voucher", voucher, append = TRUE, row.names = FALSE)
dbWriteTable(con, "promotion", promotion, append = TRUE, row.names = FALSE)
dbWriteTable(con, "product", product, append = TRUE, row.names = FALSE)
dbWriteTable(con, "order", order, append = TRUE, row.names = FALSE)
dbWriteTable(con, "sell", sell, append = TRUE, row.names = FALSE)
dbWriteTable(con, "supplier", supplier, append = TRUE, row.names = FALSE)
dbWriteTable(con, "promote", promote, append = TRUE, row.names = FALSE)
dbWriteTable(con, "stock", stock, append = TRUE, row.names = FALSE)


```

```{r}
# Read a table from the database into a data frame
customer_data <- dbReadTable(con, "customer")
ad_data <- dbReadTable(con, "ad")
warehouse_data <- dbReadTable(con, "warehouse")
promotion_data <- dbReadTable(con, "promotion")
voucher_data <- dbReadTable(con,"voucher")
product_data <- dbReadTable(con, "product")
order_data <- dbReadTable(con, "order")
sell_data <- dbReadTable(con, "sell")
supplier_data <- dbReadTable(con, "supplier")
promote_data <- dbReadTable(con, "promote")
stock_data <- dbReadTable(con, "stock")

# Print or inspect the data frames
print(customer_data)
print(ad_data)
print(warehouse_data)
print(promotion_data)
print(voucher_data)
print(product_data)
print(order_data)
print(sell_data)
print(supplier_data)
print(promote_data)
print(stock_data)


# Close the database connection
#RSQLite::dbDisconnect(my_db)
```


All the data is well-connected to the table without any errors.

-----------------------------------------------------------------
