---
title: "db_schema"
author: "DM 21"
date: "2024-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RSQLite)
```

## Connect Database
```{r connect}
my_db <- RSQLite::dbConnect(RSQLite::SQLite(),"e_commerce.db")
```

## Create entities

### Customer table
In the customer entity table, it's advisable to split the customer_name attribute into cust_first_name and cust_last_name to ensure flexibility and consistency within the database. Additionally, the cust_referral column acts as a self-referencing foreign key, linking to the customer table's primary key, cust_id. This column facilitates tracking customer referrals within the database by establishing relationships within the customer entity.
```{sql, connection = my_db}
--customer
CREATE TABLE IF NOT EXISTS customer(
  cust_id INT PRIMARY KEY,
  cust_reg_date TIMESTAMP NOT NULL DEFAULT CURRENT_TIMESTAMP,
  cust_last_name VARCHAR (100) NOT NULL,
  cust_first_name VARCHAR (100) NOT NULL,
  cust_email VARCHAR (100) UNIQUE NOT NULL,
  cust_password VARCHAR(64) NOT NULL,
  cust_phone VARCHAR (20) UNIQUE NOT NULL,
  cust_zipcode VARCHAR (10), 
  cust_street_address VARCHAR (255),
  cust_city VARCHAR (50),
  cust_county VARCHAR (10),
  membership VARCHAR(10) CHECK (membership IN ('PLATINUM', 'GOLD', 'SILVER', 'BRONZE')),
  cust_birth_date DATE,
  cust_referral INT,
  FOREIGN KEY (cust_referral) REFERENCES customer(cust_id) --self referencing
  
);
```

### Product table
The structure of product name is category_name + model_number, for example, guitar_w5x9. we split it into category_name and model_name so that we can distinguish its category and avoid 3NF, each attribute is functionally dependent only on the primary key, and there are no transitive dependencies between non-prime attributes.
```{sql, connection = my_db}
-- product
CREATE TABLE IF NOT EXISTS product(
  product_id INT PRIMARY KEY,
  category_name VARCHAR(100),
  model_name VARCHAR(100),
  selling_price INT,
  cost_price INT,
  stock_qty INT
);
```

### Supplier table
The Supplier's table shares the same address type as the customer and warehouse tables, including zipcode, street address, city, and county.
```{sql, connection = my_db}
-- supplier
CREATE TABLE IF NOT EXISTS supplier (
  s_id INT PRIMARY KEY,
  s_name VARCHAR(100),
  s_zipcode VARCHAR(10),
  s_street_address VARCHAR(255),
  s_city VARCHAR(100),
  s_county VARCHAR(20),
  s_phone VARCHAR(20) UNIQUE NOT NULL,
  s_email VARCHAR(255) UNIQUE NOT NULL
);
```

### Warehouse table
The 'warehouse' table also used the same address style with 'customer' and 'supplier' table for consistency.
```{sql, connection = my_db}
-- warehouse 
CREATE TABLE IF NOT EXISTS 'warehouse' (
  w_id INT PRIMARY KEY,
  w_name VARCHAR(100) NOT NULL,
  w_zipcode VARCHAR(10) NOT NULL,
  w_street_address VARCHAR(255) NOT NULL,
  w_city VARCHAR(100),
  w_county VARCHAR(25)
);
```

### Ad table
The 'ad' table displays basic advertisement information data along with its performance metrics, including impression, click, and action. Impressions, clicks, and actions are represented as integers, while cost and revenue allow for two decimal places to calculate the pounds.
```{sql, connection = my_db}
-- ad
CREATE TABLE IF NOT EXISTS 'ad'(
  ad_id INT PRIMARY KEY,
  media_name VARCHAR(100),
  ad_type VARCHAR(100),
  start_date DATE,
  end_date DATE,
  impression INT,
  click INT,
  cost DECIMAL(10, 2),
  action INT,
  revenue DECIMAL(10, 2)
 );
```

### Promotion table
The 'promotion_name' column is defined as a VARCHAR as it is the names of the promotion
```{sql, connection = my_db}
-- promotion
CREATE TABLE IF NOT EXISTS promotion (
 promotion_id INT PRIMARY KEY,  
 promotion_name VARCHAR (200)
); 
```

### Voucher table
The 'voucher_rate' column is defined as a float to ensure that it can accommodate percentages.
```{sql, connection = my_db}
-- voucher
CREATE TABLE IF NOT EXISTS voucher (
 voucher_code VARCHAR PRIMARY KEY,  
 voucher_rate FLOAT,
 promotion_id INT,
 FOREIGN KEY (promotion_id) REFERENCES promotion(promotion_id)
); 
```


## Create relationship 

### Order table
This table contains the most foreign keys from other entities, including 'order_id', 'cust_id', 'product_id', and 'voucher_code'. In this relationship, we have ensured that 'order_id', 'cust_id', and 'product_id' form a composite primary key.
```{sql, connection = my_db}
-- order
CREATE TABLE IF NOT EXISTS "order" (
  "order_id" INT,
  "order_date" DATE,
  "quantity" INT NOT NULL,
  "cust_id" INT,
  "product_id" INT,
  "payment_method" VARCHAR,
  "voucher_code" INT,
  PRIMARY KEY ("order_id", "cust_id", "product_id"),
  FOREIGN KEY ("cust_id") REFERENCES "customer" ("cust_id"),
  FOREIGN KEY ("product_id") REFERENCES "product" ("product_id"),
  FOREIGN KEY ("voucher_code") REFERENCES "voucher" ("voucher_code")
);
```


### Sell table
The sell table comprises two foreign keys, s_id and product_id, indicating the association between suppliers and the products they sell. It is assumed that each supplier sells distinct products, ensuring clarity and accuracy in tracking product-supplier relationships within the database.
```{sql, connection = my_db}
-- sell
CREATE TABLE IF NOT EXISTS sell(
  s_id INT,
  product_id INT,
  FOREIGN KEY (product_id) REFERENCES product(product_id),
  FOREIGN KEY (s_id) REFERENCES supplier(s_id)
);
```

### Stock table
The sku serves as the primary key of the stock table, indicating the product_id located in different warehouses.
```{sql, connection = my_db}
-- stock
CREATE TABLE IF NOT EXISTS stock(
  sku INT PRIMARY KEY,
  product_id INT,
  w_id INT,
  FOREIGN KEY (product_id) REFERENCES product(product_id),
  FOREIGN KEY (w_id) REFERENCES warehouse(w_id)
);
```

### Promote table
The 'promote' table facilitates a many-to-many relationship between products and advertisements. Each 'ad_id' can promote several 'product_id's, and conversely, each 'product_id' can be promoted by several 'ad_id's.
```{sql, connection = my_db}
-- promote

CREATE TABLE IF NOT EXISTS promote(
  product_id INT,
  ad_id INT,
  FOREIGN KEY (product_id) REFERENCES product(product_id),
  FOREIGN KEY (ad_id) REFERENCES ad (ad_id)
);
```

